Lesson 5
========================================================

### Multivariate Data
Notes: Here we will do Data Analysis using not 2 but more variables.

***

###Load Data and packages
```{r Prep work}
install.packages("reshape2")
library(ggplot2)
library(tidyr)
library(dplyr)
library(knitr)
library(reshape2)
pf <- read.csv("pseudo_facebook.tsv", sep = "\t")
```


### Third Qualitative Variable
Notes: Here we plot a boxplot for age by gender and also show the mean value for each group.

```{r Third Qualitative Variable - 1}
ggplot(aes(x = gender, y = age), data = subset(pf, !is.na(gender))) + 
  geom_boxplot() +
  geom_point(stat = "summary", fun.y = mean, shape = 3) +
  scale_y_continuous(breaks = seq(0, 113, 10))
```

Notes: Here we plot the median friend count for each gender
```{r Third Qualitative Variable - 2}
ggplot(aes(x = age, y = friend_count), 
                            data = subset(pf, !is.na(pf$gender))) +
  geom_line(aes(colour = gender), stat = "summary", fun.y = median)
```

Notes: Here we create a new data frame which has the mean and median friend count grouped by age and gender
```{r Third Qualitative Variable - 3}
pf.fc_by_age_gender <- pf %>%
  filter(!is.na(pf$gender)) %>% 
  group_by(age, gender) %>% 
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>% 
  ungroup() %>% 
  arrange(age)
```

***

### Plotting Conditional Summaries
Notes:Here we plot median friend count using the new data frame that we just created above

```{r Plotting Conditional Summaries}
ggplot(aes(y = median_friend_count, x = age, colour = gender),
       data = pf.fc_by_age_gender) +
  geom_line()
```

***

### Reshaping Data
Notes: Now we use the spread function of the tidyr package to change the columns of a our data frame.
```{r rehape using dplyr}
medianUsingDplyr <- pf.fc_by_age_gender[, c(1, 2, 4)] %>% 
  spread(gender, median_friend_count)
```

Notes: Next we will use the reshape2 package to change the data from long to wide format
```{r reshape using reshape2}
medianUsingReshape <- dcast(pf.fc_by_age_gender, age ~ gender, value.var = "median_friend_count")
```

### Wide and Long Format
Notes: The wide format of the data contains columns for different categorical variables. The long format of the data has one columnn for multiple variable values.

***

### Ratio Plot
Notes: Using the long dataframe that we created above we plot the ratio of median friend count for females and males. We also plot a horizontal line with the y intercept as 1

```{r Ratio Plot}
ggplot(aes(x = age, y = female/male), data = medianUsingReshape) +
  geom_line() +
  geom_hline(yintercept = 1, lty = 2, colour = "orange") +
  scale_y_continuous(breaks = seq(0, 3, .2)) +
  scale_x_continuous(breaks = seq(13, 120, 10))
```

***

### Third Quantitative Variable
Notes: We will add a variable called year_joined to the data frame.

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure/365)

summary(pf$year_joined)
table(pf$year_joined)
```

***

### Cut a Variable
Notes: We add another variable year joined.bucket to the data frame. This variable bins together the year_joined column in different buckets.

```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004, 2009, 2011, 2012, 2014))

table(pf$year_joined.bucket, useNA = "ifany")
```

***

### Plotting it All Together
Notes:We now plot the median friend count vs age for each of the different year_joined.bucket bins.

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count), 
       data = subset(pf, !is.na(pf$year_joined.bucket))) + 
  geom_line(aes(colour = year_joined.bucket), stat = "summary", fun.y = median) +
  scale_y_continuous(breaks = seq(0, 2000, 100)) +
  scale_x_continuous(breaks = seq(13, 120, 10))
```

***

### Plot the Grand Mean
Notes:Plot showing friend count vs age for each of the different year joined bucket bins.

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count), 
       data = subset(pf, !is.na(pf$year_joined.bucket))) +
  geom_line(aes(colour = year_joined.bucket), stat = "summary", 
            fun.y = mean) +
  geom_line(stat = "summary", fun.y = mean, lty = 2) +
  scale_y_continuous(breaks = seq(0, 2000, 100)) +
  scale_x_continuous(breaks = seq(13, 120, 5))
```

***

### Friending Rate
Notes:Calculating the friend count for the number of days active on facebook excluding the tenure of 0 days.

```{r Friending Rate}
with(subset(pf, pf$tenure > 0), summary(friend_count / tenure))
```

What is the median friend rate? - 0.2205

What is the maximum friend rate? - 417
***

### Friendships Initiated
Notes:

```{r Friendships Initiated}
pf %>% 
  filter(tenure > 0) %>% 
  ggplot(aes(x = tenure, y = friendships_initiated / tenure)) +
  geom_line(aes(colour = year_joined.bucket), stat = "summary", fun.y = mean) +
  scale_y_continuous(breaks = seq(0, 10, .5)) +
  scale_x_continuous(breaks = seq(0, 3200, 200))
```

***

### Bias-Variance Tradeoff Revisited
Notes:Here we plot the friend rate vs tenure for different bin widths. Finally we add a smoothner to the plot.

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = tenure, y = friendships_initiated / tenure), 
       data = subset(pf, pf$tenure > 0)) +
  stat_smooth(aes(colour = year_joined.bucket))
```

***
### Introducing the Yogurt Data Set
Notes:Load the data and convert id variable into a factor

```{r Load Yogurt}
yo <- read.csv("yogurt.csv")
yo$id <- as.factor(yo$id)
```

***

### Histograms Revisited
Notes:Lots of gaps in the histogram plot. Most of the yogurt bought is in the 65-70 price range.

```{r Histograms Revisited}
ggplot(aes(x = price), data = yo) +
  geom_histogram(binwidth = 1, colour = "black", fill = "orange") + 
  scale_x_continuous(breaks = seq(0, 75, 5)) +
  scale_y_continuous(breaks = seq(0, 800, 50))
```

***

### Number of Purchases
Notes:We add a column to the data frame. This new column contains the sum of all yogurt purchases for an observation.

```{r Number of Purchases}
yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(aes(x = time, y = price), data = yo) +
  geom_jitter(alpha = 1/20, colour = "orange")
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

